A Survey of Parallelization of Kalman Filters

### 2017-06-10 paper/paper.tex
aas done as it is going to get

@@

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{climate.png}
\caption{Speedup from \cite{lyster1997parallel}}
\label{fig:climate}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{mpdse.png}
\caption{System flowchart from \cite{karimipour2015extended}}
\label{fig:mpdse}
\end{figure}

@@

\subsection{Fully Distributed}
In \cite{rao1991fully} Rao presents an algorithm for a Kalmn filter across a fully distributed group of sensor nodes. In this case there is no centralized node to perform the final computation that brings the local filter results together. To make up for the lack of a central node, the authors simply broadcast the intermediate results to all the node and then every node computes the final result.

\subsection{Other sensor stuff}
\cite{spanos2005distributed,spanos2005approximate}

@@

\section{Conclusion}
Sorry I didn't finish in time. Stay tuned for updates!

In conclusion, the numerous applications of Kalman filters present varied opportunities for parallelization, but differences in the applications make it challenging to generalize a solution to multiple applications. As a result, there are a variety of creative methods used in parallel Kalman filters. These range from vectorization of the linear algebra to distributing the filter among sensor networks.

>>>>

In general there are three motivations for parallelizing an Implementation of the Kalman filter. First is for the sake of real-time performance. In many cases, particularly those for tracking, the results must be obtained in real time or else the plane, boat, or subatomic particle will be lost. Second, the number of states involved in the computation is so large that a serial implementation results in unacceptable runtime, as is the case in the climate applications. Lastly the parallelization may be inherent to the system analyzing the data such as sensor networks. In each situation the application has significant impact on the types of parallelism and computational resources available to the programmer. These differences inspire a wide variety of implementations.

@@

\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{climate.png}
\caption{Speedup from \cite{lyster1997parallel}}
\label{fig:climate}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{mpdse.png}
\caption{System flowchart from \cite{karimipour2015extended}}
\label{fig:mpdse}
\end{figure}

@@

\subsection{Fully Distributed}
In \cite{rao1991fully} Rao presents an algorithm for a Kalman filter across a fully distributed group of sensor nodes. In this case there is no centralized node to perform the final computation that brings the local filter results together. To make up for the lack of a central node, the authors simply broadcast the intermediate results to all the node and then every node computes the final result.

Spanos et al present a related system in \cite{spanos2005distributed,spanos2005approximate} that is generalizable to algorithms other than the Klamn filter. Their algorithms allow distributed node to come to a consensus at different points in the algorithm depending on the current network conditions. Their method is similar to Rao's in that it allows the nodes to work independently the n collaborate to get a final result, but achieves even greater resiliency by being flexible in when collaboration occurs. The authors show that the system can compute the filter locally, incorporating data from other sensors as it becomes available with communication occurring $n$ times per filter.

@@

While FPGA implementations provide fantastic benefits in terms of speed and power they are currently not commonly available to most systems. There has been some talk of incorporating FPGA fabric into standard chips as an adaptable accelerator, so these may become more common in the future. Also for large state-space Kalman filters the size of the hardware may become prohibitive.

\section{Conclusion}
In conclusion, the numerous applications of Kalman filters present varied opportunities for parallelization, but differences in the applications make it challenging to generalize a solution to multiple applications. As a result, there are a variety of creative methods used in parallel Kalman filters. These range from vectorization of the linear algebra to distributing the filter among sensor networks.

Potential future work includes seeking improvements in all of these areas. Notably absent from the research is the use of GPU technology; only the paper on smart grid makes good use of them. The HEP papers mention GPUs repeatedly in future work but have yet to implement anything. GPUs are challenging as the require long vectors or matrices to be effective, but these are unavailable in many of the applications. Vector registers offer a better alternative since they are a more appropriate size, but these to are increasing in length beyond what most tracking applications require. It may be possible to combine the computation for multiple filters into longer vectors that fill the entire vector to achieve the desired speedup. Additional research could be at the higher levels of the application, incorporating the Kalman filters into other parallel structures that offer more concurrency.
